---
title: "P4 Explore and Summarize Data"
author: "M.J. van der Poel"
date: "5 maart 2017"
output:
  html_document: default
  pdf_document: default
---

Red Wine Quality
========================================================

This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

The guiding question in this investigation is Which chemical properties 
influence the quality of red wines?


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(GGally)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
RW <- read.csv("wineQualityReds.csv")
```

Input variables (based on physicochemical tests):  
   1 - fixed acidity (tartaric acid - g / dm^3)  
   2 - volatile acidity (acetic acid - g / dm^3)  
   3 - citric acid (g / dm^3)  
   4 - residual sugar (g / dm^3)  
   5 - chlorides (sodium chloride - g / dm^3  
   6 - free sulfur dioxide (mg / dm^3)  
   7 - total sulfur dioxide (mg / dm^3)  
   8 - density (g / cm^3)  
   9 - pH  
   10 - sulphates (potassium sulphate - g / dm3)  
   11 - alcohol (% by volume)  
     
   Output variable (based on sensory data):  
   12 - quality (score between 0 and 10)  
   
Description of attributes:  
   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile
   (do not evaporate readily)  
   2 - volatile acidity: the amount of acetic acid in wine, which at too high 
   of levels can lead to an unpleasant, vinegar taste  
   3 - citric acid: found in small quantities, citric acid can add 'freshness' 
   and flavor to wines  
   4 - residual sugar: the amount of sugar remaining after fermentation stops, 
   it's rare to find wines with less than 1 gram/liter and wines with greater 
   than 45 grams/liter are considered sweet  
   5 - chlorides: the amount of salt in the wine  
   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between
   molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial
   growth and the oxidation of wine  
   7 - total sulfur dioxide: amount of free and bound forms of S02; in low
   concentrations, SO2 is mostly undetectable in wine, but at free SO2 
   concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine  
   8 - density: the density of water is close to that of water depending on the
   percent alcohol and sugar content  
   9 - pH: describes how acidic or basic a wine is on a scale from 0 
   (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale  
   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas 
   (S02) levels, wich acts as an antimicrobial and antioxidant  
   11 - alcohol: the percent alcohol content of the wine  

   Output variable (based on sensory data):  
   12 - quality (score between 0 and 10)  

```{r echo=FALSE, Mutate_the_Data}
# Mutate the Data
RW<-mutate(RW, bound.sulfur.dioxide=total.sulfur.dioxide-free.sulfur.dioxide)
```

Total sulfur dioxide is an addition of free and bound sulfite. 
Created a separate column to present the bound sulfur by substracting free from
the total sulfur.

# Univariate Plots Section


```{r echo=FALSE, Univariate_Plots_1}

RW$quality<-factor(RW$quality, levels=c(3,4,5,6,7,8), ordered = TRUE)

str(RW)

head(RW)

summary(RW)

```

The dataset consists of 12 variables, with 1599 observations. 
There are no missing values and the dataset is structured and tidy.
X is the unique key that identifies each wine. 
All the other fields except quality are quantitive measurements for that wine.
Quality is a qualitive field at the ordinal level. I therefore changed the
integer values into factors ranging from 3 up and until 8.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_fixed.acidity}

qplot(data=RW, x=fixed.acidity, binwidth=.2) 

```

Most wines have a fixed acidity between 4 and 12 g / dm^3 with a mean of 8,32.
The distribution is normal and a little bit right skewed.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_volatile.acidity}

qplot(data=RW, x=volatile.acidity, binwidth=.03, xlim = c(0,1.2))
```

Volatile acidity seems to behave like a bimodal distribution with most values
between 0.2 and 1.00 g / dm^3, and with a peak at around 0.4 and 0.6 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_citric.acid}
qplot(data=RW, x=citric.acid, binwidth=.04, xlim = c(0,0.8))
```

The citric acid is more or less evenly distributed between 0.02 and 0.5 g / dm^3
and then drops lineair until 0.78 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_residual.sugar}
qplot(data=RW, x=residual.sugar, binwidth=.1, xlim = c(0,10))


```

Residual sugar looks like a normal distribution between 1.0 and 3.0 but has a
very long low tail up till 15.00.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_chlorides}
qplot(data=RW, x=chlorides, binwidth=.005, xlim = c(0,0.2))
```

Chlorides is a normal distribution a little bit skewed on the right. Most values
between 0.04 and 0.13 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_free.sulfur.dioxide}
ggplot(data=RW, aes(x=free.sulfur.dioxide)) +
  geom_histogram(bins=50) 

ggplot(data=RW, aes(x=free.sulfur.dioxide)) +
  geom_histogram(bins=20) +
  scale_x_continuous(trans='log10')
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots_bound.sulfur.dioxide}
ggplot(data=RW, aes(x=bound.sulfur.dioxide)) +
  geom_histogram(bins=50) 

ggplot(data=RW, aes(x=bound.sulfur.dioxide)) +
  geom_histogram(bins=20) +
  scale_x_continuous(trans='log10')
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots_total.sulfur.dioxide}
qplot(data=RW, x=total.sulfur.dioxide, binwidth=4, xlim = c(0,175))

ggplot(data=RW, aes(x=total.sulfur.dioxide)) +
  geom_histogram(bins=20) +
  scale_x_continuous(trans='log10')
```

Free, bound and total sulfur dioxide have a normal distribution on a log 10 
scale.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_density}
qplot(data=RW, x=density, binwidth=.0002)
```

The mean and median are almost the same here. Nice normal distribution.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_pH}
qplot(data=RW, x=pH, binwidth=.02)
```

The mean and median are almost the same here.Nice normal distribution.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_sulphates}
qplot(data=RW, x=sulphates, binwidth=.02, xlim = c(0,1.5))
```

Sulphates is a normal distribution a little bit skewed on the right. Most values
between 0.4 and 1.0 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_alcohol}
qplot(data=RW, x=alcohol, binwidth=.1)



```

Alcohol percentage is between 9 and 14. It quickly builds up to a peak at 9.5
percent and then drops lineairly to zero at around 14 percent.
This is probably due to the fact that alcohol stops the yeasting proces that
is responsible for the creation of alcohol.
I wonder how the relation between alcohol and residual sugar looks, since the
alcohol is a product of the fermentation of the original sugar.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_quality, fig.height=6}

ggplot(data=RW, aes(x=quality)) + geom_bar() +  geom_text(stat='count',aes(label=..count..),vjust=-1)

```

The majority of the wines have a score 5 or 6. I wonder how the wines that
score a 3 or 8 differ from the majority.


# Univariate Analysis

### What is the structure of your dataset?

There are 1599 observations of red wines in the dataset with 12 features (fixed
acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur
dioxide, total sulfur dioxide, density, pH, sulphates, alcohol % and quality).

All variables except quality are quantitive, continuous variables. 
Quality is an ordered factor variable with the following levels from worst to
best: 3, 4, 5, 6, 7, 8.

Other observations:

* Sulfur dioxides have a normal distribution on a log10 scale
* Citric acid is more or less evenly distributed between 0.02 and 0.5 g / dm^3 
and then drops lineair until 0.78 g / dm^3.
* Alcohol percentage quickly builds from 9 percent up to a peak at 9.5 percent
and then drops lineairly to zero at around 14 percent.
* Volatile acidity seems to behave like a bimodal distribution.
* The mean and the median are almost equal for Ph and density.
* All other continuous variables have a normal distribution with some skewness.
* 82.5 % of the wines have a quality of 5 or 6. 1,75% has a quality of 3 or 8.

### What is/are the main feature(s) of interest in your dataset?


The main feature in the dataset is quality. I would like to determine Which
chemical properties influence the quality of red wines
Relation quality for every atrribute (facetwrap it or freqpolygram with colors)
Relation alcohol - residual sugar

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The relation between alcohol and residual sugar is interesting because alcohol
is formed during fermentation of sugar. The remaining sugar influences
the sweetness of the wine.
Another important taste influencer is acidity expressed in the three types of
acid(ity) and PH.
Finally the tanin determines the dryness. Tannin is not measured, but just like
free sulfur dioxide, it acts as an anti oxidant.
Lower levels of free sulfur dioxide could indicate higher (not measured) levels
of tannin.


### Did you create any new variables from existing variables in the dataset?

I created bound.sulfur.dioxide by subtracting free.sulfur.dioxide from
total.sulfur.dioxide.

### Of the features you investigated, were there any unusual distributions? \

I did not find unusual distributions and therefore made no adjustments.
I did factorize quality.


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

!!!! Facet wrap all variables over quality. Or maybe better use geom='freqpoly'and color=quality

```{r echo=FALSE, warning=FALSE, Bivariate_Plots}

cor(subset(RW, select=fixed.acidity:alcohol))

```

Correlation > 0,7 = Large, 0,5 - 0,7 is Moderate, 0,3 - 0,5 small byt meaningfull.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_ggpairs}

ggpairs (subset(RW, select=fixed.acidity:quality))

```



# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
Sites used:
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3472855/
https://www.stat.berkeley.edu/classes/s133/factors.html
