---
title: "P4 Explore and Summarize Data"
author: "M.J. van der Poel"
date: "5 maart 2017"
output:
  html_document: default
  pdf_document: default
---

Red Wine Quality
========================================================

This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

The guiding question in this investigation is Which chemical properties 
influence the quality of red wines?


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.


library(ggplot2)
library(knitr)
library(dplyr)
library(GGally)
library(corrplot)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
RW <- read.csv("wineQualityReds.csv")
```

Input variables (based on physicochemical tests):  
   1 - fixed acidity (tartaric acid - g / dm^3)  
   2 - volatile acidity (acetic acid - g / dm^3)  
   3 - citric acid (g / dm^3)  
   4 - residual sugar (g / dm^3)  
   5 - chlorides (sodium chloride - g / dm^3  
   6 - free sulfur dioxide (mg / dm^3)  
   7 - total sulfur dioxide (mg / dm^3)  
   8 - density (g / cm^3)  
   9 - pH  
   10 - sulphates (potassium sulphate - g / dm3)  
   11 - alcohol (% by volume)  
     
   Output variable (based on sensory data):  
   12 - quality (score between 0 and 10)  
   
Description of attributes:  
   1 - fixed acidity: most acids involved with wine or fixed or nonvolatile
   (do not evaporate readily)  
   2 - volatile acidity: the amount of acetic acid in wine, which at too high 
   of levels can lead to an unpleasant, vinegar taste  
   3 - citric acid: found in small quantities, citric acid can add 'freshness' 
   and flavor to wines  
   4 - residual sugar: the amount of sugar remaining after fermentation stops, 
   it's rare to find wines with less than 1 gram/liter and wines with greater 
   than 45 grams/liter are considered sweet  
   5 - chlorides: the amount of salt in the wine  
   6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between
   molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial
   growth and the oxidation of wine  
   7 - total sulfur dioxide: amount of free and bound forms of S02; in low
   concentrations, SO2 is mostly undetectable in wine, but at free SO2 
   concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine  
   8 - density: the density of water is close to that of water depending on the
   percent alcohol and sugar content  
   9 - pH: describes how acidic or basic a wine is on a scale from 0 
   (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale  
   10 - sulphates: a wine additive which can contribute to sulfur dioxide gas 
   (S02) levels, wich acts as an antimicrobial and antioxidant  
   11 - alcohol: the percent alcohol content of the wine  

   Output variable (based on sensory data):  
   12 - quality (score between 0 and 10)  

```{r echo=FALSE, Mutate_the_Data}
# Mutate the Data
RW <- subset(RW, total.sulfur.dioxide < 150 & alcohol <= 14.0)
RW<-mutate(RW, bound.sulfur.dioxide=total.sulfur.dioxide-free.sulfur.dioxide)
```

Total sulfur dioxide is an addition of free and bound sulfite. 
Created a separate column to present the bound sulfur by substracting free from
the total sulfur.

# Univariate Plots Section


```{r echo=FALSE, Univariate_Plots_1}

str(RW)

head(RW)

summary(RW)

```

The dataset consists of 12 variables, with 1599 observations. 
There are no missing values and the dataset is structured and tidy.
X is the unique key that identifies each wine. 
All the other fields except quality are quantitive measurements for that wine.
Quality is a qualitive field at the ordinal level. I therefore changed the
integer values into factors ranging from 3 up and until 8.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_fixed.acidity}

qplot(data=RW, x=fixed.acidity, binwidth=.2) 

```

Most wines have a fixed acidity between 4 and 12 g / dm^3 with a mean of 8,32.
The distribution is normal and a little bit right skewed.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_volatile.acidity}

qplot(data=RW, x=volatile.acidity, binwidth=.03, xlim = c(0,1.2))
```

Volatile acidity seems to behave like a bimodal distribution with most values
between 0.2 and 1.00 g / dm^3, and with a peak at around 0.4 and 0.6 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_citric.acid}
qplot(data=RW, x=citric.acid, binwidth=.04, xlim = c(0,0.8))
```

The citric acid is more or less evenly distributed between 0.02 and 0.5 g / dm^3
and then drops lineair until 0.78 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_residual.sugar}
qplot(data=RW, x=residual.sugar, binwidth=.1, xlim = c(0,10))


```

Residual sugar looks like a normal distribution between 1.0 and 3.0 but has a
very long low tail up till 15.00.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_chlorides}
qplot(data=RW, x=chlorides, binwidth=.005, xlim = c(0,0.2))
```

Chlorides is a normal distribution a little bit skewed on the right. Most values
between 0.04 and 0.13 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_free.sulfur.dioxide}
ggplot(data=RW, aes(x=free.sulfur.dioxide)) +
  geom_histogram(bins=50) 

ggplot(data=RW, aes(x=free.sulfur.dioxide)) +
  geom_histogram(bins=20) +
  scale_x_continuous(trans='log10')
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots_bound.sulfur.dioxide}
ggplot(data=RW, aes(x=bound.sulfur.dioxide)) +
  geom_histogram(bins=50) 

ggplot(data=RW, aes(x=bound.sulfur.dioxide)) +
  geom_histogram(bins=20) +
  scale_x_continuous(trans='log10')
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots_total.sulfur.dioxide}
qplot(data=RW, x=total.sulfur.dioxide, binwidth=4, xlim = c(0,175))

ggplot(data=RW, aes(x=total.sulfur.dioxide)) +
  geom_histogram(bins=20) +
  scale_x_continuous(trans='log10')
```

Free, bound and total sulfur dioxide have a normal distribution on a log 10 
scale.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_density}
qplot(data=RW, x=density, binwidth=.0002)
```

The mean and median are almost the same here. Nice normal distribution.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_pH}
qplot(data=RW, x=pH, binwidth=.02)
```

The mean and median are almost the same here.Nice normal distribution.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_sulphates}
qplot(data=RW, x=sulphates, binwidth=.02, xlim = c(0,1.5))
```

Sulphates is a normal distribution a little bit skewed on the right. Most values
between 0.4 and 1.0 g / dm^3.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_alcohol}
qplot(data=RW, x=alcohol, binwidth=.1)



```

Alcohol percentage is between 9 and 14. It quickly builds up to a peak at 9.5
percent and then drops lineairly to zero at around 14 percent.
This is probably due to the fact that alcohol stops the yeasting proces that
is responsible for the creation of alcohol.
I wonder how the relation between alcohol and residual sugar looks, since the
alcohol is a product of the fermentation of the original sugar.

```{r echo=FALSE, warning=FALSE, Univariate_Plots_quality, fig.height=6}

ggplot(data=RW, aes(x=quality)) + geom_bar() +  geom_text(stat='count',aes(label=..count..),vjust=-1)

```

The majority of the wines have a score 5 or 6. I wonder how the wines that
score a 3 or 8 differ from the majority.


# Univariate Analysis

### What is the structure of your dataset?

There are 1599 observations of red wines in the dataset with 12 features (fixed
acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur
dioxide, total sulfur dioxide, density, pH, sulphates, alcohol % and quality).

All variables except quality are quantitive, continuous variables. 
Quality is an ordered variable with the following levels from worst to
best: 3, 4, 5, 6, 7, 8.

Other observations:

* Sulfur dioxides have a normal distribution on a log10 scale
* Citric acid is more or less evenly distributed between 0.02 and 0.5 g / dm^3 
and then drops lineair until 0.78 g / dm^3.
* Alcohol percentage quickly builds from 9 percent up to a peak at 9.5 percent
and then drops lineairly to zero at around 14 percent.
* Volatile acidity seems to behave like a bimodal distribution.
* The mean and the median are almost equal for Ph and density.
* All other continuous variables have a normal distribution with some skewness.
* 82.5 % of the wines have a quality of 5 or 6. 1,75% has a quality of 3 or 8.

### What is/are the main feature(s) of interest in your dataset?


The main feature in the dataset is quality. I would like to determine Which
chemical properties influence the quality of red wines.
I expect that quality is at least determined by the right combination of sweetness 
(residual sugar and alcohol) and sourness (fixe and volatile acidity, citric acid
and PH).


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The relation between alcohol and residual sugar is interesting because alcohol
is formed during fermentation of sugar. The remaining sugar influences
the sweetness of the wine.
Another important taste influencer is acidity expressed in the three types of
acid(ity) and PH.
Finally the tanin determines the dryness. Tannin is not measured, but just like
free sulfur dioxide, it acts as an anti oxidant.
Lower levels of free sulfur dioxide could indicate higher (not measured) levels
of tannin.


### Did you create any new variables from existing variables in the dataset?

I created bound.sulfur.dioxide by subtracting free.sulfur.dioxide from
total.sulfur.dioxide.
At first I factorised quality with "#RW$quality<-factor(RW$quality, levels=c(3,4,5,6,7,8), ordered = TRUE)", but changed it back because it gave errors with the cor function when I wanted to correlate quality with the other fields.

### Of the features you investigated, were there any unusual distributions? \

Total sulphur dioxide has two ourliers that I removed. Alcohol had one outlier
that I removed.


# Bivariate Plots Section

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_1}

My_cor <- cor(subset(RW, select=fixed.acidity:bound.sulfur.dioxide))

My_cor

corrplot(My_cor, method="circle", type="lower")
corrplot(My_cor, method="number", type="lower", number.cex=0.75)

```

I used the following guidelines for correlation:
If > 0,7 = Large, 0,7 - 0,5 is Moderate, 0,5 - 0,3 small but meaningfull.

Based on this I see a moderate positive correlation between:  
* fixed acidity and citric acid  
* fixed acidity and density  

I see a moderate negative correlation between:  
* fixed acidity and PH  
* volatile acidity and citric acid  
* citric acid and PH  
* alcohol and density  

I see meaningfull positive correlation between:  
* citric acid and density  
* citric acid and sulphates  
* residual sugar and density  
* chlorides and sulphates  
* free sulfur dioxide and bound sulfur dioxode  
* quality and alcohol  

I see meaningfull negative correlation between:  
* quality and volatile acidity  
* PH and density  

I on purpose did not mention the correlation between bound and free sulfur
dioxide on the one hand and total dioxide on the other hand.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_2}

ggplot(data = RW, aes(x=citric.acid, y=fixed.acidity)) + geom_point(alpha=0.2, colour="blue") + geom_smooth(method = "gam") + ggtitle ("Citric Acid vs Fixed Acidity") +
  xlab("Citric.acid in g / dm^3") + ylab("Fixed Acidity in g / dm^3")

ggplot(data = RW, aes(x=citric.acid, y=volatile.acidity)) + geom_point(alpha=0.2, colour="green") + geom_smooth(method = "gam") + ggtitle ("Citric Acid vs Volatile Acidity") +
  xlab("Citric.acid in g / dm^3") + ylab("Volatile Acidity in g / dm^3")

ggplot(data = RW, aes(x=fixed.acidity, y=volatile.acidity)) + geom_point(alpha=0.2, colour="red") + geom_smooth(method = "gam") + ggtitle ("Fixed Acidity vs Volatile Acidity") +
  xlab("Fixed Acidity in g / dm^3") + ylab("Volatile Acidity in g / dm^3")


```

While citric acid and fixed acidity are strongly positively correlated, 
volatile acidity is negatively correlated with the other two.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_3}

RW$quality <- factor(RW$quality,levels=c("3","4","5","6","7","8"))

# Boxplot of fixed.acidity by quality 
ggplot(RW, aes(quality,fixed.acidity)) +
  geom_boxplot() +
  ggtitle("Fixed Acidity per Quality level") + xlab("Quality") + ylab("Fixed Acidity in g / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 

# Boxplot of volatile.acidity by quality 
ggplot(RW, aes(quality,volatile.acidity)) +
  geom_boxplot() +
  ggtitle("Volatile Acidity per Quality level") + xlab("Quality") + ylab("Volatile Acidity in g / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 

# Boxplot of citric.acid by quality 
ggplot(RW, aes(quality,citric.acid)) +
  geom_boxplot() +
  ggtitle("Citric acid per Quality level") + xlab("Quality") + ylab("Citric acid in g / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 


```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_4}
ggplot(data = RW, aes(x=pH, y=fixed.acidity)) + geom_point(alpha=0.2, colour="blue") + geom_smooth(method = "gam") + ggtitle ("pH vs Fixed Acidity") +
  xlab("pH") + ylab("Fixed Acidity in g / dm^3")

ggplot(data = RW, aes(x=pH, y=citric.acid)) + geom_point(alpha=0.2, colour="green") + geom_smooth(method = "gam") + ggtitle ("pH vs Citric Acid") +
  xlab("pH") + ylab("Citric.acid in g / dm^3")

ggplot(data = RW, aes(x=pH, y=volatile.acidity)) + geom_point(alpha=0.2, colour="red") + geom_smooth(method = "gam") + ggtitle ("pH vs Volatile Acidity") +
  xlab("pH") + ylab("Volatile Acidity in g / dm^3")

```

pH is very strongly influenced by fixed acidity (clear direction, close to the
smooth line),a little bit less strong (data points more spread out) by citric 
acid, and the least by volatile acidity. More fixed acidity and citric acid add
to the overal acidity taste, more volatile acidity to less overal acidity taste.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_10}

# Boxplot of PH by quality 
ggplot(RW, aes(quality,pH)) +
  geom_boxplot() +
  ggtitle("pH per Quality level") + xlab("Quality") + ylab("pH") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 

```

From the correlation information and the boxplots it is clear that acidity, acid
and PH are strongly correlated to each other. But, is it the volatile acidity
and the citric acid that are correlated to quality. The box plots for these two
show that higher levels of citric acid and lower levels of volatile acid on
average lead to a higher quality. What is also noticable, that for quality
levels 7 and 8, there is hardly any difference. Maybe to discriminate between a
7 and 8, other factors play a more important role.
For fixed acidity I don't see significant differences on level 5 and 6 and no
trend taking the other levels into account. I suppose this is because the the
majority of measurements is on level 5 and 6.
A higher fixed acidity does lead to a lower PH. Higher quality wines tend to
lower PH levels. 

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_5}

# Boxplot of residual.sugar by quality 
ggplot(RW, aes(quality,residual.sugar)) +
  geom_boxplot() +
  ggtitle("Residual Sugar per Quality level") + xlab("Quality") + ylab("Residual Sugar in g / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 

```

I cannot discover any relation between residual sugar and quality. The median
for the various quality levels is not moving in a certain direction and there
are lots of outliers for most quality levels.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_6}

# Boxplot of free.sulfur.dioxide by quality 
ggplot(RW, aes(quality,free.sulfur.dioxide)) +
  geom_boxplot() +
  ggtitle("Free sulfur dioxide per Quality level") + xlab("Quality") + ylab("Free sulfur dioxide in mg / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 

```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_7}

# Boxplot of bound.sulfur.dioxide by quality 
ggplot(RW, aes(quality,bound.sulfur.dioxide)) +
  geom_boxplot() +
  ggtitle("Bound sulfur dioxide per Quality level") + xlab("Quality") + ylab("Bound sulfur dioxide in mg / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_8}

# Boxplot of total.sulfur.dioxide by quality 
ggplot(RW, aes(quality,total.sulfur.dioxide)) +
  geom_boxplot() +
  ggtitle("Total sulfur dioxide per Quality level") + xlab("Quality") + ylab("Total sulfur dioxide in mg / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 
```

Free and bound (and therefore total) sulfur dioxide does not show a linear 
relationship with quality, but a normal-like distribution: low levels of SO2
for low and high quality wines and higher levels of SO2 for the average 
quality wines. I wonder if in the right combination with other variables like 
alcohol, sulphates and/or PH, SO2 can improve or worsen a wine.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_9}

# Boxplot of density by quality 
ggplot(RW, aes(quality,density)) +
  geom_boxplot() +
  ggtitle("Density per Quality level") + xlab("Quality") + ylab("Density in g / cm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 


ggplot(data = RW, aes(x=density, y=fixed.acidity)) + geom_point(alpha=0.2, colour="blue") + geom_smooth(method = "gam") + ggtitle ("Density vs Fixed Acidity") +
  xlab("Density in g / cm^3") + ylab("Fixed Acidity in g / dm^3")
ggplot(data = RW, aes(x=density, y=citric.acid)) + geom_point(alpha=0.2, colour="red") + geom_smooth(method = "gam") + ggtitle ("Density vs Citric Acid") +
  xlab("Density in g / cm^3") + ylab("Citric Acid in g / dm^3")
ggplot(data = RW, aes(x=density, y=residual.sugar)) + geom_point(alpha=0.2, colour="green") + geom_smooth(method = "gam") + ggtitle ("Density vs Residual Sugar") +
  xlab("Density in g / cm^3") + ylab("Residual Sugar in g / dm^3")
ggplot(data = RW, aes(x=density, y=alcohol)) + geom_point(alpha=0.2, colour="purple") + geom_smooth(method = "gam") + ggtitle ("Density vs Alcohol") +
  xlab("Density in g / cm^3") + ylab("Alcohol in % by volume")

```

Although density has a correlation with quality, I believe this is caused by the
effect that primarily alcohol and fixed acidity have on density. 
Alcohol lowers the density, residual sugar, citric acid and fixed acidity
raises the density.


```{r echo=FALSE, warning=FALSE, Bivariate_Plots_11}

# Boxplot of sulphates by quality 
ggplot(RW, aes(quality,sulphates)) +
  geom_boxplot() +
  ggtitle("Sulphates per Quality level") + xlab("Quality") + ylab("Sulphates in g / dm^3") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 
```

Higher quality wines have more sulphates on average. Between quality 7 and 8
sulphates are not a discriminating factor anymore.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots_12}


# Boxplot of alcohol by quality 
ggplot(RW, aes(quality,alcohol)) +
  geom_boxplot() +
  ggtitle("Alcohol % per Quality level") + xlab("Quality") + ylab("Alcohol %") +
  stat_summary(fun.y=mean, geom="line", aes(group=1))  +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) 

```

Alcohol has the strongest and positive correlation with quality.
I notice a strange dip in what seems a linear relation between the two at
quality level 5, both for mean and median. Is that because the SO2 amount at
quality level 5 is very high compared to the other quality levels?
You could also look at it as follows from quality level 5 till 8 there is a
lineair raise in quality with the alcohol %. The same form level 3 till 4.
But from 3 till 4 the raise is less steep and the mean and median % is higher
then 5.  


# Bivariate Analysis

### How did the feature of interest vary with other features in the dataset?

The feature of interest is quality. I found a number of correlations:  

1.  Citric Acid: Correlation +0.23. Higher quality wines tend to have
    higher citric acid levels. Between quality levels 7 and 8 I don't see much
    difference.    
2.  Volatile Acidity: Correlation -0.39. Higher quality wines tend to have
    lower volitale acidity levels. Again, between quality levels 7 and 8 I don't
    see much difference, except that the standard deviation for level 8 is smaller.    
3.  Bound Sulphur Dioxide: Correlation -0.21. I calculated the bound SO2 by
    removing the free SO2 from the total SO2. The bound SO2 has the strongest
    correlation with quality of the three. They all have a normal distribution,
    meaning that both low and high quality wines have on average low levels of SO2
    and medium quality wines have high levels of SO2.   
4.  Sulphates: Correlation +0.25. Higher quality wines tend to have higher
    sulphates levels. Between quality levels 7 and 8 the median level is the same,
    but the standard deviation on level 8 is much smaller.  
5.  Alcohol %: Highest correlation with quality of +0.48. Higher quality wines
    on average have higher alcohol levels, but there is a dip on quality level 5.
    I wonder if this dip can be explained by the high SO2 levels on quality level 5.
    I checked quality level 5 for the other variables and PH shows a simular,
    though difficult to see, dip. I want to examine alcohol, bound SO2, PH and
    quality with one another.  


### Did you observe any interesting relationships between the other features?

I expected that Citric acid and fixed acidity would have a negative correlation
with ph, I didn't expect volatile acidity to have a positive correlation.
I think, that in this case it is relative, meaning, they all contribute to the
acidity levels, but volatile acidity contributes much less, and therefore
contibutes to Ph levels on the high end of the range.

Density has a relation with four variables: Alcohol lowers the density, 
residual sugar, citric acid and fixed acidity raises the density.

### What was the strongest relationship you found?

The strongest relation I found was between total SO2 and bound SO2. But since I
derived bound SO2 from total SO2, this is no real surprise.

The next strongest relation I found between PH and fixed acidity (-0.68). 

The strongest relation I found with the feature of interest quality was with
alcohol (+0.48).

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

ggplot(aes(y = bound.sulfur.dioxide, x = alcohol, group=quality, colour=quality), data = RW) +   
  geom_line() + facet_wrap(~quality, ncol = 3) +
  geom_smooth(method = "gam")

ggplot(aes(y = bound.sulfur.dioxide, x = alcohol, group=quality, colour=quality), data = subset(RW, quality=="3" | quality=="4" | quality=="8")) +   
  geom_point(aes(colour = quality)) 
  

ggplot(aes(x = pH, y = alcohol, group=quality, colour=quality), data = RW) +   
  geom_line() + facet_wrap(~quality, ncol = 3) +
  geom_smooth(method = "gam")

ggplot(aes(x = pH, y = alcohol, group=quality, colour=quality), data = subset(RW, quality=="3" | quality=="4" | quality=="8")) +   
  geom_point(aes(colour = quality))

ggplot(aes(x = pH, y = bound.sulfur.dioxide, group=quality, colour=quality), data = RW) +   
  geom_line() + facet_wrap(~quality, ncol = 3) +
  geom_smooth(method = "gam")

ggplot(aes(x = pH, y = bound.sulfur.dioxide, group=quality, colour=quality), data = subset(RW, quality=="3" | quality=="4" | quality=="8")) +   
  geom_point(aes(colour = quality))


```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
Sites used:
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3472855/
https://www.stat.berkeley.edu/classes/s133/factors.html
